---
title: Tutorial on Using FasterRisk inside R
# output: html_document
author: Jiachang Liu
date: "Last compiled on `r format(Sys.time(), '%B %d, %Y')`"
output:
    md_document:
        variant: markdown_github
        toc: true
        number_sections: true
        includes: 
            in_header: header.md
---

<!-- Rscript -e "rmarkdown::render('header.Rmd'); rmarkdown::render('README.Rmd')" -->

# Installation

## Install the reticulate Package
```{r, results='hide', message=FALSE, warning=FALSE}
install.packages("reticulate", repos = "http://cran.us.r-project.org")
library(reticulate)
```

## Create a Python Virtual Environment
```{r, results='hide', message=FALSE}
version <- "3.9.12"
install_python(version)
virtualenv_create(envname="FasterRisk-environment", version = version)
use_virtualenv("FasterRisk-environment")
```

## Install the FasterRisk Package from PyPI
```{r, results='hide', message=FALSE, warning=FALSE}
py_install("fasterrisk", pip=TRUE, envname="FasterRisk-environment")
```

# Preparation before Training
```{r, warning=FALSE}
fasterrisk <- import("fasterrisk")
```


## Download Sample Data
```{r}
train_data_file_path <- "../tests/adult_train_data.csv"
test_data_file_path <- "../tests/adult_test_data.csv"

if (!file.exists(train_data_file_path)){
    fasterrisk$utils$download_file_from_google_drive('1nuWn0QVG8tk3AN4I4f3abWLcFEP3WPec', train_data_file_path)
}

if (!file.exists(train_data_file_path)){
    fasterrisk$utils$download_file_from_google_drive('1TyBO02LiGfHbatPWU4nzc8AndtIF-7WH', test_data_file_path)
}
```

## Read Sample Data
```{r}
np <- import("numpy", convert=FALSE)
train_df <- read.csv(train_data_file_path)
train_data <- data.matrix(train_df)
X_train <- np$array(train_data[, 2:ncol(train_data)])
y_train <- np$array(train_data[, 1], dtype=np$int)

test_df <- read.csv(test_data_file_path)
test_data <- data.matrix(test_df)
X_test <- np$array(test_data[, 2:ncol(test_data)])
y_test <- np$array(test_data[, 1], dtype=np$int)
```

# Training the Model

## Create A Model Class
```{r}
sparsity <- as.integer(5)
parent_size <- as.integer(10)

RiskScoreOptimizer_m <- fasterrisk$fasterrisk$RiskScoreOptimizer(X = X_train, y = y_train, k = sparsity, parent_size = parent_size)
```

## Train the Model
```{r}
start_time <- Sys.time()
RiskScoreOptimizer_m$optimize()
sprintf("Optimization takes %f seconds.", Sys.time() - start_time)
```

# Get Risk Score Models

## Get Solutions from the Trained Model Class
```{r}
solutions = RiskScoreOptimizer_m$get_models()
multipliers = solutions[1][[1]]
sparseDiversePool_beta0_integer = solutions[2][[1]]
sparseDiversePool_betas_integer = solutions[3][[1]]
sprintf("We generate %d risk score models from the sparse diverse pool", length(multipliers))
```

## Access the First Risk Score Model
```{r}
model_index = 1 # first model
multiplier = multipliers[model_index]
intercept = sparseDiversePool_beta0_integer[model_index]
coefficients = np$array(sparseDiversePool_betas_integer[model_index, ])
```

## Use the First Risk Score Model to Do Prediction
```{r}
RiskScoreClassifier_m = fasterrisk$fasterrisk$RiskScoreClassifier(multiplier, intercept, coefficients)
```

```{r}
y_test_pred = RiskScoreClassifier_m$predict(X_test)
print("y_test are predicted to be (first 10 values):")
y_test_pred[1:10]
```

```{r}
y_test_pred_prob = RiskScoreClassifier_m$predict_prob(X_test)
print("The risk probabilities of having y_test to be +1 are (first 10 values):")
y_test_pred_prob[1:10]
```

## Print the First Model Card
```{r}
X_featureNames = list(colnames(train_df)[-1])[[1]]

RiskScoreClassifier_m$reset_featureNames(X_featureNames)
tmp_str = py_capture_output(RiskScoreClassifier_m$print_model_card(), type = c("stdout", "stderr"))
cat(tmp_str)
```

## Print Top 10 Model Cards from the Pool and Their Performance Metrics
```{r}
num_models = min(10, length(multipliers))

for (model_index in 1:num_models){
    multiplier = multipliers[model_index]
    intercept = sparseDiversePool_beta0_integer[model_index]
    coefficients = np$array(sparseDiversePool_betas_integer[model_index, ])

    RiskScoreClassifier_m = fasterrisk$fasterrisk$RiskScoreClassifier(multiplier, intercept, coefficients)
    RiskScoreClassifier_m$reset_featureNames(X_featureNames)
    # RiskScoreClassifier_m$print_model_card()
    tmp_str = py_capture_output(RiskScoreClassifier_m$print_model_card(), type = c("stdout", "stderr"))
    cat(tmp_str)

    train_loss = RiskScoreClassifier_m$compute_logisticLoss(X_train, y_train)
    train_results = RiskScoreClassifier_m$get_acc_and_auc(X_train, y_train)
    train_acc = train_results[1][[1]]
    train_auc = train_results[2][[1]]
    test_results = RiskScoreClassifier_m$get_acc_and_auc(X_test, y_test)
    test_acc = test_results[1][[1]]
    test_auc = test_results[2][[1]]

    tmp_str = sprintf("The logistic loss on the training set is %f", train_loss)
    print(tmp_str)
    tmp_str = sprintf("The training accuracy and AUC are %f and %f", train_acc*100, train_auc)
    print(tmp_str)
    tmp_str = sprintf("The test accuracy and AUC are are %f and %f", test_acc*100, test_auc)
    print(tmp_str)
    cat("\n")
}
```